---
title: "How Income Changes Affect Personally Experienced Inflation"
author: "Owen Jetton"
date: "5/25/2022"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
```

## Motivation

Much is hidden in the Bureau of Labor Statistic’s monthly report of the Consumer Price Index that does not yield a holistic view of inflation’s effect on consumers and individuals in the U.S. economy.

Literature review:   
- Higher income individuals experience lower inflation through to access to product variety and innovation (Jaravel, 2017)   
- Lower income individuals experienced higher inflation for the costs of living during the Great Recession (Argente & Lee, 2015)    

No causal research... yet?

## Causal Question of Interest

**If someone experiences an increase to their income, how does this cause their personally experienced inflation to change?**

Calculation of personally experienced inflation:

$$\pi_{t}^i = \frac{\sum_{l}{\theta_{l,t}^iP_{l,t}} - \sum_{l}{\theta_{l,t-1}^iP_{l,t-1}}} {\sum_{l}{\theta_{l,t-1}^iP_{l,t-1}}}$$
As vectors:
$$\pi_{t}^i = [\theta_{t}^{i\prime}P_{t} - \theta_{t-1}^{i\prime}P_{t-1}] [\theta_{t-1}^{i\prime}P_{t-1}]^{-1}$$


## Potential Sources of Bias

```{r echo = F}

p_load(tidyverse, ggdag)  

coordinates = list(
  x = c(N = 0, I = 1, B = 2, E = 3, P = 1.5, L = 2), # Node name = x coordinate for that node
  y = c(N = 1, I = 1, B = 1, E = 1, P = 0, L = 2) # Node name = y coordinate for that node
)

dag = dagify(I ~ N + L,
                 B ~ P + I,
                 E ~ L + B,
                 P ~ I,
                 coords = coordinates,
                 labels = c(N = "Nominal \n Income",
                            I = "Purchasing \n Power",
                            B = "Budget \n Allocations",
                            E = "Personally \n Experienced \n Inflation ",
                            P = "Preferences",
                            L = "Price Level"))


ggdag(dag, 
      text = FALSE,
      use_labels = "label") +
  theme_dag()

```


Preferences & income | MPC | Income & Types of Goods | Locations & inflation shocks | Seasonal Shocks | Methods used 


## Ideal Experiment

* Subjects: a large group of people with identical income streams, living in same location.
* Track their consumption of individual goods perfectly
* Randomly assign portion of group to receive a sizeable permanent pay increase
* Continue to track the consumption of the treatment & control groups
* Calculate personally experienced inflation & use diff-in-diff for point estimate


## Practical Research Design (quasi-experimental)

* Large city planning a pay increase for teachers in one district but not another (not unheard of)

* With their permission, track their bank records before and after the pay raise (difficult but not impossible to obtain)

* Calculate personally experienced inflation & use diff-in-diff for point estimate: $\beta_4$ from:

$$\pi_t^i = \beta_0 + \beta_1 D_i + \beta_2 T_t + \beta_3 CPI_t + \beta_4 D_i T_t + \epsilon_{it} $$


## Possible Results: Positive Effect

```{r echo = F}

set.seed(11131998)

data = tibble(
  x = -12:12,
  control = 0.5 + rnorm(25, 0, 0.2),
  y_pos = if_else(x < 0, 0.5 + rnorm(25, 0, 0.2), 1 + rnorm(25, 0, 0.2)),
  y_null = 0.5 + rnorm(25, 0, 0.2),
  y_neg = if_else(x < 0, 0.5 + rnorm(25, 0, 0.2), 0 + rnorm(25, 0, 0.2))
)

avg_b = data %>% 
          filter(x < 0) %>% 
          summarize(c_avg = mean(control),
                    y_pos = mean(y_pos),
                    y_null = mean(y_null),
                    y_neg = mean(y_neg))
avg_a = data %>% 
          filter(x >= 0) %>% 
          summarize(c_avg = mean(control),
                    y_pos = mean(y_pos),
                    y_null = mean(y_null),
                    y_neg = mean(y_neg))

# graph positive effect
ggplot(data, aes(x = x)) +
  geom_vline(xintercept = 0, linetype = "longdash") +
  labs(x = "Months from Income Change",
       y = "Personally Experienced Inflation (%)") +
  geom_segment(aes(x = -12, xend = 0, y = avg_b$c_avg, yend = avg_b$c_avg), color = "blue") +
  geom_segment(aes(x = -12, xend = 0, y = avg_b$y_pos, yend = avg_b$y_pos), color = "red") +
  geom_segment(aes(x = 0, xend = 12, y = avg_a$c_avg, yend = avg_a$c_avg), color = "blue") +
  geom_segment(aes(x = 0, xend = 12, y = avg_a$y_pos, yend = avg_a$y_pos), color = "red") +
  geom_point(aes(y = control), color = "blue", size = 3) +
  geom_point(aes(y = y_pos), color = "red", size = 3)
```

## Possible Results: Null Effect

```{r echo = F}
 
# graph null effect
ggplot(data, aes(x = x)) +
  geom_vline(xintercept = 0, linetype = "longdash") +
  labs(x = "Months from Income Change",
       y = "Personally Experienced Inflation (%)") +
  geom_segment(aes(x = -12, xend = 0, y = avg_b$c_avg, yend = avg_b$c_avg), color = "blue") +
  geom_segment(aes(x = 0, xend = 12, y = avg_a$c_avg, yend = avg_a$c_avg), color = "blue") +
  geom_segment(aes(x = -12, xend = 0, y = avg_b$y_null, yend = avg_b$y_null), color = "red") +
  geom_segment(aes(x = 0, xend = 12, y = avg_a$y_null, yend = avg_a$y_null), color = "red") +
  geom_point(aes(y = control), color = "blue", size = 3) +
  geom_point(aes(y = y_null), color = "red", size = 3)
```

## Possible Results: Negative Effect

```{r echo = F}
# graph negative effect 
ggplot(data, aes(x = x)) +
  geom_vline(xintercept = 0, linetype = "longdash") +
  labs(x = "Months from Income Change",
       y = "Personally Experienced Inflation (%)") +
  geom_segment(aes(x = -12, xend = 0, y = avg_b$c_avg, yend = avg_b$c_avg), color = "blue") +
  geom_segment(aes(x = 0, xend = 12, y = avg_a$c_avg, yend = avg_a$c_avg), color = "blue") +
  geom_segment(aes(x = -12, xend = 0, y = avg_b$y_neg, yend = avg_b$y_neg), color = "red") +
  geom_segment(aes(x = 0, xend = 12, y = avg_a$y_neg, yend = avg_a$y_neg), color = "red") +
  geom_point(aes(y = control), color = "blue", size = 3) +
  geom_point(aes(y = y_neg), color = "red", size = 3)


```

## Sources & Extra

[Argente, David and Munseob Lee, “*Cost of Living Inequality during the Great Recession*,” 2015](https://economicdynamics.org/meetpapers/2015/paper_1372.pdf)

[Jaravel, Xavier, “*The unequal gains from product innovations: Evidence from the US retail sector*,” 2017](https://equitablegrowth.org/wp-content/uploads/2017/03/031417-WP-unequal-gains-from-product-innovations.pdf)

[What's Your Rate of Inflation? NYT](https://www.nytimes.com/interactive/2022/05/08/business/economy/inflation-calculator.html)


